A Data-Driven Framework for Neural Field Modeling

### 2011-01-13 ltx/first_order_framework/LaTeX_Manuscript.tex
minor edits-took out line on minimal assumptions
moved assumptions section as ken and parham suggested

@@

\subsection{State and Parameter Estimation}\label{StateAndParameterEstimationSection} In this section, we describe the procedure for estimating the states, $\mathbf{x}_t$, the connectivity kernel parameters, $\boldsymbol \theta$, and the synaptic dynamics, $\xi$. The estimation process is a two part iterative algorithm, consisting of a state estimation step followed by a parameter estimation step. Once initialized, at each iteration the sequence of estimated state vectors is used to update the parameter set. The resulting parameters are then used when estimating the new state vector sequence for the next iteration. The procedure stops when the parameter estimates converge. The algorithm is initialized using a bounded random state vector sequence, guaranteeing that the initial estimated parameter set forms a stable kernel.

@@

\subsection{Modeling Assumptions}
\dean{(comment 1 and 9 from reviewer 2:) It is an implicit assumption that the neural field model (in equation~\ref{DiscreteTimeModel}) provides an apt description of the cortical dynamics recorded from the subject. Although models of this form are capable of describing a variety of cortical dynamics, there is without doubt a mismatch between the cortex and the model. Nevertheless, there is a sufficient volume of interesting results from theoretical studies using the WC field equations that warrants the data assimilation framework. }

\dean{An advantage in using a lumped-parameter field model over a more detailed mathematical description is that the myriad of parameters that might influence excitability, such as specific ion concentrations, are considered to be lumped into the connectivity kernel coefficients that effectively describe the net system gain. In this way, a minimal number of assumptions are made. A major challenge in data assimilation for the mass action of the brain is to make models sufficiently detailed, such that the parameters are meaningful, but simple enough to yield clear insights that can be related to theoretical studies (from both neuroscience and engineering perspectives).}

\dean{The framework assumes that the parameters of the model (and brain) are stationary over the estimation period. This is a reasonable assumption when applying the framework to an evoked-potential experiment over a short time period in a controlled environment. However, it is well known that factors that influence excitability are non-stationary over longer time periods (see \citet{Ullah2010}). Therefore, the estimation framework must be altered if it is to be applied to data over extended periods of time. A possible way to achieve this is to use a more sophisticated model that accounts for longer-term time dependancies (higher temporal order). Alternatively, the parameters of the model could be considered time-varying, where changes in excitability would result in different connectivity kernel parameters.}

 % A possible way to achieve this would be to employ window-based estimation or fixed-lag, where the estimation is performed on successive windows of data (possibly overlapping windows). The window length will need to be sufficiently short such that the parameters are approximately stationary of the estimation duration, and sufficiently long to achieve a good estimation result. If this trade-off can not be successfully achieved then the level of detail in the model must be increased. A real-time implementation can also be achieved using the UKF \parham{and an online recursive LS estimator (see \citet{Ljung1999} for example)}. }

\subsection{Extensions to the Framework}
In this study, an assumption was made that the chosen kernel basis functions can exactly represent the actual connectivity kernel. In future work, we plan to relax this assumption \dean{and show how the support of the connectivity kernel can also be inferred from data}. In addition, we have assumed that the neural field is isotropic and homogeneous. It is generally accepted that the intracortical connectivity structure can be accurately modeled as being isotropic. However, the assumption of homogeneity is quite strong as the size of the neural field becomes larger. Future work should focus on developing estimation algorithms where assumptions of homogeneity can be relaxed. In addition, natural extensions of this framework are the inclusion of a post-synaptic response kernel with a finite rise and decay time, and a finite distant-dependant action potential propagation velocity into the parametric neural field equations.

\mike{(Reviewer 2, comment 9:)A further extension of the framework is to allow for the algorithm to learn online. The estimation framework that we have presented is inherently a batch algorithm designed to operate after data has been collected. \parham{A simple real-time implementation can be achieved using the UKF  and an online recursive  LS estimator \cite{Ljung1999}, more sophisticated approaches include the incremental EM algorithm \cite{Neal1998}, and the stepwise EM \citep{Sato2000,Cappe2009}.}}

@@

% \clearpage
% \newpage
% \begin{figure*}[!ht]
% \begin{center}
% \includegraphics{fig1.pdf} 
% \end{center}
% \caption{}
% \label{fig:Figure1}
% \end{figure*}
% \clearpage
% \newpage
% \begin{figure*}[!ht]
% \begin{center}
% \includegraphics{fig2.pdf} 
% \end{center}
% \caption{}
% \label{fig:Figure2}
% \end{figure*}
% \clearpage
% \newpage
% \begin{figure*}[!ht]
% \begin{center}
% \includegraphics{fig3.pdf}
% \end{center}
% \caption{} 
% \label{fig:Figure3}
% \end{figure*}
% \clearpage
% \newpage
% \begin{figure*}[!ht]
% \begin{center}
% \includegraphics{fig4.pdf}
% \end{center}
% \caption{}
% \label{fig:Figure4}
% \end{figure*}
% \clearpage
% \newpage
% \begin{figure*}[!ht] 
% \begin{center}
% \includegraphics{fig5.pdf}	
% \end{center}
% \caption{}
% \label{fig:Figure5}
% \end{figure*}
% \clearpage
% \newpage
% \begin{figure*}[!ht]
% \begin{center}
% \includegraphics{fig6.pdf}
% \end{center}
% \caption{}
% \label{fig:Figure6}
% \end{figure*}
% \clearpage
% \newpage
% \begin{figure*}[!ht]
% \begin{center}
% \includegraphics{fig7.pdf}
% \end{center}
% \caption{} 
% \label{fig:Figure7}
% \end{figure*}
% \clearpage
% \newpage
% \begin{figure*}[!ht]
% \begin{center}
% \includegraphics{fig8.pdf} 
% \end{center}
% \caption{} 
% \label{fig:Figure8}
% \end{figure*}
% \clearpage
% \newpage
% \begin{figure*}[!ht]
% \begin{center}
% \includegraphics{fig9.pdf}
% \end{center}
% \caption{}
% \label{fig:Figure9}
% \end{figure*}
% \clearpage
% \newpage
% \begin{figure*}[!ht]
% \begin{center}
% \includegraphics{fig10.pdf}
% \end{center}
% \caption{}
% \label{fig:Figure10}
% \end{figure*}
% \clearpage
% \newpage
% \begin{figure*}[!ht]
% \begin{center}
% \includegraphics{fig11.pdf}
% \end{center}
% \caption{}
% \label{fig:Figure11}
% \end{figure*}
% \clearpage
% \newpage
% \begin{figure*}[!ht]
% \begin{center}
% \includegraphics{fig12.pdf}
% \end{center}
% \caption{}
% \label{fig:Figure12}
% \end{figure*}
% \clearpage
% \newpage
% \begin{figure*}[!ht]
% \begin{center}
% \includegraphics{fig13.pdf}
% \end{center}
% \caption{}
% \label{fig:SensorAndSource}
% \end{figure*}
% \clearpage
% \newpage
% \begin{figure*}[!ht]
% \begin{center}
% \includegraphics{fig14.pdf}
% \end{center}
% \caption{}
% \label{fig:Ch3}
% \end{figure*}
\clearpage
\newpage
\appendix
\section{Time Discretization of the Neural Field Model}
\label{App:Time Discretization} To form the integro-difference equation neural field model, a time discretization must be performed. To do this, we used a one-step Euler method, where equation~\ref{FinalFormContinuous} can be approximated by 
\begin{align}
	\label{Euler Approximation}\frac{v\left( \mathbf{r},t+T_s \right) - v\left( \mathbf{r},t\right)}{T_s} &=-\zeta v\left( \mathbf{r},t \right) \nonumber \\ 
&+ \int_\Omega {w\left( \mathbf{r}-\mathbf{r}' \right)f\left( {v\left( \mathbf{r}',t \right)} \right)d\mathbf{r}'}. 
\end{align}
For clarity, we shall index time points in the discrete time form of the model using the subscript $t$ and the next time point as $t+1$. Rearranging equation~\ref{Euler Approximation}, we get 
\begin{equation}
	\label{Euler Approximation2} v_{t+1}\left( \mathbf{r}\right) = v_t\left( \mathbf{r}\right) -T_s \zeta v_t\left( \mathbf{r}\right) + T_s \int_\Omega {w\left( \mathbf{r}-\mathbf{r}' \right)f\left( {v_t\left( \mathbf{r}'\right)} \right)d\mathbf{r}'}. 
\end{equation}
Now making simplifications, the discrete time form of the model is 
\begin{equation}
	\label{Discrete Time Model1}v_{t+1}\left(\mathbf{r}\right) = \xi v_t\left(\mathbf{r}\right) + T_s \int_\Omega { w\left(\mathbf{r}-\mathbf{r}'\right) f\left(v_t\left(\mathbf{r}'\right)\right) d\mathbf{r}'}, 
\end{equation}
where $\xi = 1 - T_s \zeta$.

\parham{(Reviewer1, Additional comments 2:) The analysis of the error introduced by Euler discretization method can be found in \citet{Butcher2008}. Here we only study the stability of the discretized model, we show that if $0 < T_s < \tau$, where the synaptic time constant $\tau = 1/\zeta$,  the discretized IDE is stable. To do this, the model is written as the series
\begin{align}
	v_1(\mathbf{r}) &= \xi v_0(\mathbf{r}) + T_s g_0(\mathbf{r}) \\
	v_2(\mathbf{r})	&= \xi^2 v_0(\mathbf{r}) + T_s \left(\xi g_0(\mathbf{r}) + g_1(\mathbf{r}) \right) \\
	\vdots \, \, \, \, \, &= \, \, \, \, \, \vdots \nonumber \\
	v_{t+1}(\mathbf r) &= \xi^{t+1} v_0\left(\mathbf{r} \right) + T_s \sum_{i=0}^t \xi^i g_{t-i}(\mathbf{r}), \label{eq:fieldmodelseries}
\end{align}
where
\begin{equation}
	g_{t}(\mathbf{r}) = \int_\Omega { w\left(\mathbf{r}-\mathbf{r}'\right) f\left(v_t\left(\mathbf{r}'\right)\right) d\mathbf{r}'}.
\end{equation}
Since  $0 \leq f(v_t(\mathbf r)) \leq 1$ and the kernel is semi-compactly supported, $g_t(\mathbf r)$ is always bounded . To assess the stability of the discrete time system we assume that $g_t(\mathbf{r})$ is maximally saturated over the entire field for all time, which we denote as $g_{max}$. Substituting this into equation~\ref{eq:fieldmodelseries} yields
\begin{align}\label{vrecursiveinlimit}
	v_{t+1}(\mathbf r) &= \xi^{t+1}v_0(\mathbf r)+ g_{max} T_s \sum_{i=0}^t \xi^i \nonumber \\
	&=\xi^{t+1}v_0(\mathbf r) + \frac{ g_{max} T_s(1-\xi^{t+1})}{1-\xi}.
\end{align}
To study the asymptotic behaviour of the system we consider the limit as $t$ approaches infinity. Noting $0 < \xi <1$  we use a property of the geometric series to simplify equation \eqref{vrecursiveinlimit} to
\begin{equation}\label{simplifiedvrecursiveinlimit}
	v_{t+1}(\mathbf r)=\xi^{t+1}v_0(\mathbf r)+\frac{g_{max} T_s}{1-\xi}.
\end{equation}
Substituting $\xi=1-\zeta T_s$ into equation \eqref{simplifiedvrecursiveinlimit} gives
\begin{equation}
	v_{t+1}(\mathbf r)=(1-\zeta T_s)^{t+1}v_0(\mathbf r)+\frac{g_{max}}{\zeta}.
\end{equation}
From this equation it can be seen that if $0 < T_s < \tau$, the system is stable. The sampling period was set to $T_s = 0.001 s$ and the time constant was set to $\tau = 0.01 s^{-1}$. Therefore, the discretization preserved the stability properties of the system.}

@@

\section{Inner Product of Two \emph{n}-Dimensional Gaussians}\label{App:InnerProdOfGaussians}
The derivation in this section was used for calculating the matrix $\mathbf \Gamma$ from equation~\ref{eq:DefGamma}. Consider two Gaussian basis functions
\begin{equation}\label{eq:n_dimensional_Gaussian1}
 \varphi_i(\mathbf r)=\mathrm{exp}\left({-\frac{1}{\sigma_i^2} (\mathbf r-\boldsymbol \mu_i)^\top(\mathbf r-\boldsymbol \mu_i})\right)
\end{equation}
and 
\begin{equation}\label{eq:n_dimensional_Gaussian2}
\varphi_j(\mathbf r)=\mathrm{exp}\left({-\frac{1}{\sigma_j^2} (\mathbf r-\boldsymbol \mu_j)^\top(\mathbf r-\boldsymbol \mu_j})\right).
\end{equation}

>>>>

\dean{(comment 1 and 9 from reviewer 2:) It is an implicit assumption that the neural field model (in equation~\ref{DiscreteTimeModel}) provides an apt description of the cortical dynamics recorded from a specific subject. Although models of this form are capable of describing a variety of cortical dynamics, there will be without doubt a mismatch between the cortex and the model. Nevertheless, there is a sufficient volume of interesting results from theoretical studies using the WC field equations that warrants the data assimilation framework. An advantage in using a lumped-parameter field model over a more detailed mathematical description is that the myriad of parameters that might influence excitability, such as specific ion concentrations, are considered to be lumped into parameters (connectivity kernel coefficients) that effectively describe the net system gain. Therefore, there are less parameters to estimate. A major challenge in model-based data analysis for the mass action of the brain is to make models sufficiently detailed, such that the parameters are meaningful, but simple enough to yield clear insights that can be related to theoretical studies (from both neuroscience and engineering perspectives).}

@@

\subsection{State and Parameter Estimation}\label{StateAndParameterEstimationSection} In this section, we describe the procedure for estimating the states, $\mathbf{x}_t$, the connectivity kernel parameters, $\boldsymbol \theta$, and the synaptic dynamics, $\xi$. \dean{The framework assumes that the parameters of the model (and brain) are stationary over the estimation period. This is a reasonable assumption when applying the framework to an evoked-potential experiment over a short time period in a controlled environment.}

The estimation process is a two part iterative algorithm, consisting of a state estimation step followed by a parameter estimation step. Once initialized, at each iteration the sequence of estimated state vectors is used to update the parameter set. The resulting parameters are then used when estimating the new state vector sequence for the next iteration. The procedure stops when the parameter estimates converge. The algorithm is initialized using a bounded random state vector sequence, guaranteeing that the initial estimated parameter set forms a stable kernel.

@@

\subsection{Extensions to the Framework}
In this study, an assumption was made that the chosen kernel basis functions can exactly represent the actual connectivity kernel. In future work, we plan to relax this assumption \dean{and show how the support of the connectivity kernel can also be inferred from data}. In addition, we have assumed that the neural field is isotropic and homogeneous. It is generally accepted that the intracortical connectivity structure can be accurately modeled as being isotropic. However, the assumption of homogeneity is quite strong as the size of the neural field becomes larger. Future work should focus on developing estimation algorithms where assumptions of homogeneity can be relaxed. Other natural extensions of this framework are the inclusion of a post-synaptic response kernel with a finite rise and decay time, and a finite distant-dependant action potential propagation velocity into the parametric neural field equations.

\dean{The assumption was made that the parameters of the model are stationary over the estimation period. However, it is well known that factors that influence excitability are non-stationary over longer time periods (see \citet{Ullah2010}). Therefore, the estimation framework must be altered, to account for time-varying parameters, if it is to be applied to data over extended periods of time. A possible way to achieve this is to use a more sophisticated model that accounts for longer-term time dependancies (higher temporal order). Alternatively, the parameters of the model could be considered time-varying, where changes in excitability would result in different connectivity kernel parameters.}

\mike{(Reviewer 2, comment 9:) A further extension of the framework is to allow for the algorithm to learn online. The estimation framework that we have presented is inherently a batch algorithm designed to operate after data has been collected. \parham{A simple real-time implementation can be achieved using the UKF and an online recursive LS estimator \cite{Ljung1999}. Other more sophisticated approaches that may be used in a real-time estimation scheme include the incremental EM algorithm \cite{Neal1998}, and the stepwise EM \citep{Sato2000,Cappe2009}.}}

@@

% \clearpage
% \newpage
% \begin{figure*}[!ht]
% \begin{center}
% \includegraphics{fig1.pdf} 
% \end{center}
% \caption{}
% \label{fig:Figure1}
% \end{figure*}
% \clearpage
% \newpage
% \begin{figure*}[!ht]
% \begin{center}
% \includegraphics{fig2.pdf} 
% \end{center}
% \caption{}
% \label{fig:Figure2}
% \end{figure*}
% \clearpage
% \newpage
% \begin{figure*}[!ht]
% \begin{center}
% \includegraphics{fig3.pdf}
% \end{center}
% \caption{} 
% \label{fig:Figure3}
% \end{figure*}
% \clearpage
% \newpage
% \begin{figure*}[!ht]
% \begin{center}
% \includegraphics{fig4.pdf}
% \end{center}
% \caption{}
% \label{fig:Figure4}
% \end{figure*}
% \clearpage
% \newpage
% \begin{figure*}[!ht] 
% \begin{center}
% \includegraphics{fig5.pdf}	
% \end{center}
% \caption{}
% \label{fig:Figure5}
% \end{figure*}
% \clearpage
% \newpage
% \begin{figure*}[!ht]
% \begin{center}
% \includegraphics{fig6.pdf}
% \end{center}
% \caption{}
% \label{fig:Figure6}
% \end{figure*}
% \clearpage
% \newpage
% \begin{figure*}[!ht]
% \begin{center}
% \includegraphics{fig7.pdf}
% \end{center}
% \caption{} 
% \label{fig:Figure7}
% \end{figure*}
% \clearpage
% \newpage
% \begin{figure*}[!ht]
% \begin{center}
% \includegraphics{fig8.pdf} 
% \end{center}
% \caption{} 
% \label{fig:Figure8}
% \end{figure*}
% \clearpage
% \newpage
% \begin{figure*}[!ht]
% \begin{center}
% \includegraphics{fig9.pdf}
% \end{center}
% \caption{}
% \label{fig:Figure9}
% \end{figure*}
% \clearpage
% \newpage
% \begin{figure*}[!ht]
% \begin{center}
% \includegraphics{fig10.pdf}
% \end{center}
% \caption{}
% \label{fig:Figure10}
% \end{figure*}
% \clearpage
% \newpage
% \begin{figure*}[!ht]
% \begin{center}
% \includegraphics{fig11.pdf}
% \end{center}
% \caption{}
% \label{fig:Figure11}
% \end{figure*}
% \clearpage
% \newpage
% \begin{figure*}[!ht]
% \begin{center}
% \includegraphics{fig12.pdf}
% \end{center}
% \caption{}
% \label{fig:Figure12}
% \end{figure*}
% \clearpage
% \newpage
% \begin{figure*}[!ht]
% \begin{center}
% \includegraphics{fig13.pdf}
% \end{center}
% \caption{}
% \label{fig:SensorAndSource}
% \end{figure*}
% \clearpage
% \newpage
% \begin{figure*}[!ht]
% \begin{center}
% \includegraphics{fig14.pdf}
% \end{center}
% \caption{}
% \label{fig:Ch3}
% \end{figure*}
\clearpage
\newpage
\appendix
\section{Time Discretization of the Neural Field Model}
\label{App:Time Discretization} This appendix provides a description of the discretization of the neural field model to form the integro-difference equation (IDE) and provides a proof for the stability of the discrete-time equations. To perform the discretization, we used a one-step Euler method, where equation~\ref{FinalFormContinuous} can be approximated by 
\begin{align}
	\label{Euler Approximation}\frac{v\left( \mathbf{r},t+T_s \right) - v\left( \mathbf{r},t\right)}{T_s} &=-\zeta v\left( \mathbf{r},t \right) \nonumber \\ 
&+ \int_\Omega {w\left( \mathbf{r}-\mathbf{r}' \right)f\left( {v\left( \mathbf{r}',t \right)} \right)d\mathbf{r}'}. 
\end{align}
For clarity, we shall index time points in the discrete time form of the model using the subscript $t$ and the next time point as $t+1$. Rearranging equation~\ref{Euler Approximation}, we get 
\begin{equation}
	\label{Euler Approximation2} v_{t+1}\left( \mathbf{r}\right) = v_t\left( \mathbf{r}\right) -T_s \zeta v_t\left( \mathbf{r}\right) + T_s \int_\Omega {w\left( \mathbf{r}-\mathbf{r}' \right)f\left( {v_t\left( \mathbf{r}'\right)} \right)d\mathbf{r}'}. 
\end{equation}
Now making simplifications, the discrete time form of the model is 
\begin{equation}
	\label{Discrete Time Model1}v_{t+1}\left(\mathbf{r}\right) = \xi v_t\left(\mathbf{r}\right) + T_s \int_\Omega { w\left(\mathbf{r}-\mathbf{r}'\right) f\left(v_t\left(\mathbf{r}'\right)\right) d\mathbf{r}'}, 
\end{equation}
where $\xi = 1 - T_s \zeta$.

\parham{(Reviewer1, Additional comments 2:) If $0 < T_s < \tau$, where the synaptic time constant $\tau = 1/\zeta$, the IDE field model is stable. To show this, the model is written as the series
\begin{align}
	v_1(\mathbf{r}) &= \xi v_0(\mathbf{r}) + T_s g_0(\mathbf{r}) \\
	v_2(\mathbf{r})	&= \xi^2 v_0(\mathbf{r}) + T_s \left(\xi g_0(\mathbf{r}) + g_1(\mathbf{r}) \right) \\
	\vdots \, \, \, \, \, &= \, \, \, \, \, \vdots \nonumber \\
	v_{t+1}(\mathbf r) &= \xi^{t+1} v_0\left(\mathbf{r} \right) + T_s \sum_{i=0}^t \xi^i g_{t-i}(\mathbf{r}), \label{eq:fieldmodelseries}
\end{align}
where
\begin{equation}
	g_{t}(\mathbf{r}) = \int_\Omega { w\left(\mathbf{r}-\mathbf{r}'\right) f\left(v_t\left(\mathbf{r}'\right)\right) d\mathbf{r}'}.
\end{equation}
Since  $0 \leq f(v_t(\mathbf r)) \leq 1$ and the kernel is semi-compactly supported, $g_t(\mathbf r)$ is always bounded . To assess the stability of the system we assume that $g_t(\mathbf{r})$ is maximally saturated over the entire field for all time, which we denote as $g_{max}$. Substituting this into equation~\ref{eq:fieldmodelseries} yields
\begin{align}\label{vrecursiveinlimit}
	v_{t+1}(\mathbf r) &= \xi^{t+1}v_0(\mathbf r)+ g_{max} T_s \sum_{i=0}^t \xi^i \nonumber \\
	&=\xi^{t+1}v_0(\mathbf r) + \frac{ g_{max} T_s(1-\xi^{t+1})}{1-\xi}.
\end{align}
To study the asymptotic behaviour of the system we consider the limit as $t$ approaches infinity. Since $0 < \xi <1$  we use a property of the geometric series to simplify equation \eqref{vrecursiveinlimit} to
\begin{equation}\label{simplifiedvrecursiveinlimit}
	v_{t+1}(\mathbf r)=\xi^{t+1}v_0(\mathbf r)+\frac{g_{max} T_s}{1-\xi}.
\end{equation}
Substituting $\xi=1-\zeta T_s$ into equation \eqref{simplifiedvrecursiveinlimit} gives
\begin{align}
	v_{t+1}(\mathbf r) &= (1-\zeta T_s)^{t+1}v_0(\mathbf r)+\frac{g_{max}}{\zeta} \\
	&= \left(1-\frac{T_s}{\tau}\right)^{t+1}v_0(\mathbf r) + \tau g_{max}
\end{align}
From this equation it can be seen that if $0 < T_s < \tau$, the system is stable. The sampling period was set to $T_s = 0.001 s$ and the time constant was set to $\tau = 0.01 s^{-1}$. Therefore, the discretization preserved the stability properties of the system. Further analysis of the error introduced by Euler discretization method can be found in \citet{Butcher2008}. }

@@

\section{Inner Product of Two \emph{n}-Dimensional Gaussians}\label{App:InnerProdOfGaussians}
The derivation in this appendix was used for calculating the matrix $\mathbf \Gamma$ from equation~\ref{eq:DefGamma}. Consider two Gaussian basis functions
\begin{equation}\label{eq:n_dimensional_Gaussian1}
 \varphi_i(\mathbf r)=\mathrm{exp}\left({-\frac{1}{\sigma_i^2} (\mathbf r-\boldsymbol \mu_i)^\top(\mathbf r-\boldsymbol \mu_i})\right)
\end{equation}
and 
\begin{equation}\label{eq:n_dimensional_Gaussian2}
\varphi_j(\mathbf r)=\mathrm{exp}\left({-\frac{1}{\sigma_j^2} (\mathbf r-\boldsymbol \mu_j)^\top(\mathbf r-\boldsymbol \mu_j})\right).
\end{equation}
